---
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
    theme: spacelab           # flatly spacelab sandstone cerulean
    code_download: true
---

```{r docSetup, warning = FALSE, message = FALSE}
#### Record start time
startTime <- Sys.time()

library('DT')

source(here::here("scripts","DESeq_functions.R"))

species_data <- load_species(params$species)

bs <- load_biospyder_new(params$biospyder_dbs, species_data$temposeq_manifest)


```


# `r params$project_title` - Extra stats {-}

Date report generated: `r format(Sys.time(), '%d %B, %Y')`  
Report prepared for: `r params$project_name`


***

This report contains detailed tables of DEG data.

```{r load_facet_data, include=FALSE}
source(here::here("scripts","load-facet-data.R"))
```


This table shows the significant DEGs (passing all filtering criteria) ordered by their absolute fold change. Use the search function to find your feature of interest or sort by one of the columns. You can limit to a single contrast if desired.

```{r 'topFeatures', results = 'asis', warning = FALSE}
searchURL <- "http://www.ncbi.nlm.nih.gov/gene/?term="
## Add search url if appropriate
res.df.dt <- allResults %>%
  dplyr::filter(abs(linearFoldChange) > params$linear_fc_filter) %>%
  dplyr::filter(padj < params$alpha) %>%
  arrange(-abs(linearFoldChange))


descriptions <- AnnotationDbi::select(get(species_data$orgdb), columns = c("ENSEMBL", "GENENAME"), keys = res.df.dt$Ensembl_Gene_ID, keytype="ENSEMBL")
colnames(descriptions) <- c("Ensembl_Gene_ID","description")
res.df.dt <- res.df.dt %>% left_join(descriptions)


if (!is.null(searchURL)) {
    res.df.dt$Gene <- paste0('<a href="',
                             searchURL,
                             res.df.dt$Ensembl_Gene_ID,
                             '" rel="noopener noreferrer" target="_blank">',
                             res.df.dt$Ensembl_Gene_ID,
                             '<br/>',
                             res.df.dt$Gene_Symbol,
                             '</a>')
    res.df.dt <- res.df.dt %>% dplyr::relocate(Gene)
}

res.df.dt[, 'padj'] <- format(res.df.dt[, 'padj'],
                              scientific = TRUE,
                              digits = digits)
res.df.dt[, 'pvalue'] <- format(res.df.dt[, 'pvalue'],
                                scientific = TRUE,
                                digits = digits)
res.df.dt <- res.df.dt %>%
  dplyr::select(-c(Gene_Symbol, Ensembl_Gene_ID))



DT::datatable(res.df.dt,
          options = list(pagingType = 'full_numbers',
                         pageLength = 20,
                         scrollX = '100%',
                         dom = 'Bfrtip',
                         buttons = c('copy',
                                     'csv',
                                     'excel',
                                     'pdf',
                                     'print',
                                     'colvis'),
                         columnDefs = list(list(visible = FALSE, targets = c(1, 2, 4, 5)))),
          escape = FALSE,
          extensions = 'Buttons',
          rownames = FALSE,
          filter = "top",
          ) %>% 
  DT::formatRound(which(!colnames(res.df.dt) %in% c('pvalue',
                                                    'padj',
                                                    'Feature',
                                                    'contrast',
                                                    'description',
                                                    'Gene',
                                                    'Ensembl_Gene_ID',
                                                    'Probe_Name')),
                  digits)


```




# Session Info

## Date the report was generated.

```{r reproducibility1, echo = FALSE}
## Date the report was generated
Sys.time()
```

## Version of R-ODAF Repository

The git hash for the commit of the [R-ODAF repository](`r system("git remote get-url origin", intern = T)`) used in this analysis is `r system("git rev-parse HEAD", intern = T)`.

## Parameters Used

From the list elements in each params${variable} used to generate this report.

```{r paramsList, echo = FALSE}
df <- as.data.frame(unlist(params))
names(df) <- "Parameter Value"
knitr::kable(as.data.frame(df), format = "markdown")
```

## Wallclock time spent generating the report.

```{r reproducibility2, echo = FALSE}
## Processing time in seconds
totalTime <- diff(c(startTime, Sys.time()))
round(totalTime, digits = 3)
```

## `R` session information.

```{r reproducibility3, echo = FALSE}
## Session info
options(width = 120)
session_info()
```

## Pandoc version used: `r rmarkdown::pandoc_version()`.