---
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
    theme: spacelab           # flatly spacelab sandstone cerulean
    code_download: true
---


```{r docSetup, warning = FALSE, message = FALSE}
#### Record start time
startTime <- Sys.time()

library('clusterProfiler')
library('enrichplot')

source(here::here("scripts","DESeq_functions.R"))

species_data <- load_species(params$species)

bs <- load_biospyder_new(params$biospyder_dbs, species_data$temposeq_manifest)


```


# `r params$project_title` - GO and pathway analysis {-}

Date report generated: `r format(Sys.time(), '%d %B, %Y')`  
Report prepared for: `r params$project_name`


***

This report contains Gene Ontology and pathway analysis.

```{r load_facet_data, include=FALSE}
source(here::here("scripts","load-facet-data.R"))
```



# GO Enrichment Analysis {.tabset .tabset-fade}

This section performs GO enrichment on **all DEGs passing filters**. The background set of genes is those that were identified in this sequencing experiment. The [clusterProfiler](https://bioconductor.org/packages/release/bioc/html/clusterProfiler.html) package is used for this analysis.

```{r cluster-profiler, warning = FALSE, message = FALSE, collapse = TRUE}

id_table_entrez <- params$bs$biospyder %>%
  dplyr::select(!!sym(params$biomart_filter), Gene_Symbol, Entrez_ID)



probeIDDEGs <- data.frame(genes=significantResults[[params$biomart_filter]])
names(probeIDDEGs)[names(probeIDDEGs) == 'genes'] <- params$biomart_filter

DEGs <- dplyr::left_join(probeIDDEGs, id_table_entrez)
entrezDEGs <- as.character(DEGs$Entrez_ID)
entrezDEGs <- entrezDEGs[!is.na(entrezDEGs)] %>% unique()

background <- dplyr::left_join(data.frame(genes = row.names(assay(dds))),
                               id_table_entrez,
                               by = c("genes" = params$biomart_filter)) %>%
  dplyr::filter(!is.na(Entrez_ID)) %>%
  dplyr::pull(Entrez_ID)
background <- as.character(background)

KeyType <- "ENTREZID"
```

```{r enrich_go, warning = FALSE, message = FALSE, collapse = TRUE, eval=F}
enrich_go_bp <- enrichGO(gene = entrezDEGs,
                         universe = background,
                         OrgDb = species_data$orgdb,
                         keyType = KeyType,
                         readable = T,
                         ont = "BP",
                         pAdjustMethod = "BH",
                         pvalueCutoff = 0.01,
                         qvalueCutoff = 0.05)


enrich_go_mf <- enrichGO(gene = entrezDEGs,
                         universe = background,
                         OrgDb = species_data$orgdb,
                         keyType = KeyType,
                         readable = T,
                         ont = "MF",
                         pAdjustMethod = "BH",
                         pvalueCutoff = 0.01,
                         qvalueCutoff = 0.05)

enrich_go_cc <- enrichGO(gene = entrezDEGs,
                         universe = background,
                         OrgDb = species_data$orgdb,
                         keyType = KeyType,
                         ont = "CC",
                         pAdjustMethod = "BH",
                         pvalueCutoff = 0.01,
                         qvalueCutoff = 0.05)

############ TODO DELETE THIS FOR TESTING ONLY!!! #############################

enrich_go_bp@result$p.adjust[1:10] <- rep(0.001,10)
enrich_go_bp@result$qvalue[1:10] <- rep(0.001,10)

enrich_go_mf@result$p.adjust[1:10] <- rep(0.001,10)
enrich_go_mf@result$qvalue[1:10] <- rep(0.001,10)

enrich_go_cc@result$p.adjust[1:5] <- rep(0.001,5)
enrich_go_cc@result$qvalue[1:5] <- rep(0.001,5)

##############################################################################


# TODO GO table output?
# write.table(as.data.frame(enrich_go_bp),
#             file = file.path(paths$pathway_analysis, "Enriched.GO_terms.Biological_Processes.txt"))
# write.table(as.data.frame(enrich_go_mf),
#             file = file.path(paths$pathway_analysis, "Enriched.GO_terms.Molecular_Functions.txt"))
# write.table(as.data.frame(enrich_go_cc),
#             file = file.path(paths$pathway_analysis, "Enriched.GO_terms.Cellular_Component.txt"))


# get fold-changes

DEGs_full <- dplyr::left_join(significantResults,
                              id_table_entrez,
                              by = params$biomart_filter) %>%
  distinct()
foldChanges <- DEGs_full %>% dplyr::pull(log2FoldChange)
names(foldChanges) <- DEGs_full %>% dplyr::pull(Entrez_ID)
foldChanges <- foldChanges %>% sort() %>% rev()
foldChanges <- foldChanges[!is.na(names(foldChanges))]


```

## Biological processes


```{r plot_GO_analysis_bp, warning = FALSE, message = FALSE, fig.width = 12, collapse = TRUE, eval=F}
if (!is.null(enrich_go_bp)) {
  if (nrow(enrich_go_bp@result %>% filter(p.adjust < 0.01 & qvalue < 0.05)) > 3) {
    plot1 <- clusterProfiler::dotplot(enrich_go_bp, font.size = 9, showCategory = 20)  +
      theme(axis.text.y = element_text(angle = 0)) +
      scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 30))

    simplified_bp <- simplify(enrich_go_bp, cutoff = 0.7, by = "p.adjust", select_fun = min)
    # Max size of gene set... useful for visualizing networks
    simplified_bp_for_network <- simplified_bp %>% DOSE::gsfilter(by = "GSSize", max = 400)
    # I believe there is a bug in the current version of DOSE or clusterProfiler that prevents this from working
    # Throws this error on plotting: Error in graph.data.frame(x, directed = FALSE) : the data frame should contain at least two columns
    simplified_bp_filt <- simplified_bp %>%
      filter(p.adjust < .05, qvalue < 0.05) %>%
      mutate(GeneRatio = DOSE::parse_ratio(GeneRatio)) %>%
      arrange(desc(GeneRatio))

    if (nrow(as.data.frame(simplified_bp_filt)) < 10) {
        show_num <- simplified_bp_filt$Description # Top N less than 10
      } else {
        show_num <- simplified_bp_filt$Description[1:10] # Top 10
      }

    plot2 <- cnetplot(simplified_bp,
                      foldChange = foldChanges,
                      showCategory = show_num,
                      cex_label_gene = 0.3)
    
    plot3 <- upsetplot(enrich_go_bp)
    plot4 <- heatplot(simplified_bp, foldChange = foldChanges, showCategory = show_num) +
      scale_y_discrete(labels = function(x) stringr::str_trunc(x, width = 50))
    
    print(plot1)
    print(plot2)
    print(plot3)
    print(plot4)
  } else {
    num_terms <- nrow(enrich_go_bp@result %>% filter(p.adjust < 0.01 & qvalue < 0.05))
    cat(paste0(num_terms," enriched terms found. This is not enough to report in the results."))
  }
} else { cat("No significantly enriched terms using criteria selected") }
```


## Molecular Functions

```{r plot_GO_analysis_mf, warning = FALSE, message = FALSE, fig.width = 12, collapse = TRUE, eval=F}
if (!is.null(enrich_go_mf)) {
  
  if (nrow(enrich_go_mf@result %>% filter(p.adjust < 0.01 & qvalue < 0.05)) > 3) {
    plot1 <- clusterProfiler::dotplot(enrich_go_mf, font.size = 9)  +
      theme(axis.text.y = element_text(angle = 0)) +
      scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 30))
    
    simplified_mf <- simplify(enrich_go_mf, cutoff = 0.7, by = "p.adjust", select_fun = min)
    # Max size of gene set... useful for visualizing networks
    simplified_mf_for_network <- simplified_mf %>% DOSE::gsfilter(by = "GSSize", max = 400)
    # I believe there is a bug in the current version of DOSE or clusterProfiler that prevents this from working
    # Throws this error on plotting: Error in graph.data.frame(x, directed = FALSE) : the data frame should contain at least two columns
    simplified_mf_filt <- simplified_mf %>%
      filter(p.adjust < .05, qvalue < 0.05) %>%
      mutate(GeneRatio = DOSE::parse_ratio(GeneRatio)) %>%
      arrange(desc(GeneRatio))
    
    if (nrow(as.data.frame(simplified_mf)) < 10) {
        show_num <- simplified_mf_filt$Description # Top N less than 10
      } else {
        show_num <- simplified_mf_filt$Description[1:10] # Top 10
      }
    
    if (nrow(simplified_mf) > 3) {
      plot2 <- cnetplot(simplified_mf,
                        foldChange = foldChanges,
                        showCategory = show_num,
                        cex_label_gene = 0.3)
    }
    
    enrich_go_mf_upset <- enrich_go_mf
    enrich_go_mf_upset@result$Description <- str_trunc(enrich_go_mf_upset@result$Description, width = 50)
    plot3 <- upsetplot(enrich_go_mf_upset)
    plot4 <- enrichplot::heatplot(simplified_mf, foldChange = foldChanges, showCategory = show_num) +
      scale_y_discrete(labels = function(x) stringr::str_trunc(x, width = 50))
    
    print(plot1)
    if (nrow(simplified_mf) > 3) {print(plot2)} else {print("Network plot not shown due to a limited number of enriched terms.")}
    print(plot3)
    print(plot4)
  } else {
    num_terms <- nrow(enrich_go_mf@result %>% filter(p.adjust < 0.01 & qvalue < 0.05))
    cat(paste0(num_terms," enriched terms found. This is not enough to report in the results."))
  }
} else {cat("No significantly enriched terms using criteria selected") }
```

## Cellular Component

```{r plot_GO_analysis_cc, warning = FALSE, message = FALSE, fig.width = 12, collapse = TRUE, eval=F}
if (!is.null(enrich_go_cc)) {
  if (nrow(enrich_go_cc@result %>% filter(p.adjust < 0.01 & qvalue < 0.05)) > 3) {
    plot1 <- clusterProfiler::dotplot(enrich_go_cc, font.size = 9)  +
      theme(axis.text.y = element_text(angle = 0)) +
      scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 30))
    
    simplified_cc <- simplify(enrich_go_cc, cutoff = 0.7, by = "p.adjust", select_fun = min)
    # Max size of gene set... useful for visualizing networks
    simplified_cc_for_network <- simplified_cc %>% DOSE::gsfilter(by = "GSSize", max = 400)
    # I believe there is a bug in the current version of DOSE or clusterProfiler that prevents this from working
    # Throws this error on plotting: Error in graph.data.frame(x, directed = FALSE) : the data frame should contain at least two columns
    simplified_cc_filt <- simplified_cc %>%
      filter(p.adjust < .05, qvalue < 0.05) %>%
      mutate(GeneRatio = DOSE::parse_ratio(GeneRatio)) %>%
      arrange(desc(GeneRatio))

    if (nrow(as.data.frame(simplified_cc_filt)) < 10) {
      show_num <- simplified_cc_filt$Description # Top N less than 10
      } else {
        show_num <- simplified_cc_filt$Description[1:10] # Top 10
      }
    
    plot2 <- cnetplot(simplified_cc,
                      foldChange = foldChanges,
                      showCategory = show_num,
                      cex_label_gene = 0.3)
    
    enrich_go_cc_upset <- enrich_go_cc
    enrich_go_cc_upset@result$Description <- str_trunc(enrich_go_cc_upset@result$Description, width = 50)
    plot3 <- upsetplot(enrich_go_cc_upset)
    plot4 <- enrichplot::heatplot(simplified_cc, foldChange = foldChanges, showCategory = show_num) +
      scale_y_discrete(labels = function(x) stringr::str_trunc(x, width = 50))
    
    print(plot1)
    print(plot2)
    print(plot3)
    print(plot4)
  } else {
    num_terms <- nrow(enrich_go_cc@result %>% filter(p.adjust < 0.01 & qvalue < 0.05))
    cat(paste0(num_terms," enriched terms found. This is not enough to report in the results."))
  }
} else {cat("No significantly enriched terms using criteria selected")}
```


# Pathway Analysis using Enriched WikiPathways and GSEA (KEGG)

## WikiPathways: Enrichment based on all DEGs

This section uses clusterProfiler to detect enriched gene sets using [WikiPathways](https://www.wikipathways.org/) data.  

```{r 'pathway_analysis', collapse=TRUE, eval=params$run_pathway_analysis, include=params$run_pathway_analysis, warning = FALSE, message = FALSE}
entrez_IDs <- as.character(id_table_entrez$Entrez_ID)
entrezDEGs <- dplyr::left_join(significantResults,
                               id_table_entrez,
                               by = params$biomart_filter)
entrezDEGs <- as.vector(entrezDEGs$Entrez_ID)

#Wiki <- rWikiPathways::downloadPathwayArchive(organism=species_sci, format = "gmt", destpath = paths$DEG_output, date="20210310")
#wp2gene <- clusterProfiler::read.gmt(file.path(paths$DEG_output,Wiki))
wp2gene <- clusterProfiler::read.gmt(file.path(paths$wikipathways,params$wikipathways_filename))
wp2gene <- wp2gene %>% tidyr::separate(1, c("name","version","wpid","org"), "%")
wpid2gene <- wp2gene %>% dplyr::select(wpid,gene) #TERM2GENE
wpid2name <- wp2gene %>% dplyr::select(wpid,name) #TERM2NAME
wpid2gene
wpid2name
matches <- wpid2gene %>% filter(gene %in% entrezDEGs)

if (nrow(matches) > 1) {
  DEG_pathways <- clusterProfiler::enricher(
    entrezDEGs,
    universe = entrez_IDs,
    pAdjustMethod = "fdr",
    pvalueCutoff = 0.05, #p.adjust cutoff
    TERM2GENE = wpid2gene,
    TERM2NAME = wpid2name)
  DEG_pathways <- DOSE::setReadable(DEG_pathways, species_data$orgdb, keyType = "ENTREZID")
  ###################### TODO FIX THIS CHANTGED FOR TESTING ######################
  DEG_pathways@result$p.adjust[1:10] <- rep(0.001,10)
  DEG_pathways@result$qvalue[1:10] <- rep(0.001,10)
  ###################### TODO FIX THIS CHANTGED FOR TESTING ######################
  if (nrow(DEG_pathways@result %>% filter(p.adjust < 0.01 & qvalue < 0.05)) > 1) {
    asdf <- clusterProfiler::dotplot(DEG_pathways, showCategory = 20, font.size = 9)  +
      theme(axis.text.y = element_text(angle = 0)) +
      scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 30))
    print(asdf)
    write.table(as.data.frame(DEG_pathways),
                file = file.path(paths$pathway_analysis, "Enriched.WikiPathways.txt"))
  } else { cat("No significantly enriched terms using criteria selected") }
}
```


## WikiPathways: Enrichment by DEGs within each contrast {.tabset .tabset-fade}

This section uses clusterProfiler `enricher()` to detect enriched gene sets using [WikiPathways](https://www.wikipathways.org/) data. Only contrasts with results are shown here.  

```{r 'pathway_analysis_by_contrast_make_plots', collapse=TRUE, eval=params$run_pathway_analysis, include=params$run_pathway_analysis, warning = FALSE, message = FALSE}
plotList <- list()
contrasts <- unique(significantResults$contrast)
for (i in seq_along(contrasts) ){
  entrezDEGs <- dplyr::left_join(significantResults %>% filter(contrast == contrasts[i]),
                                 id_table_entrez,
                                 by = params$biomart_filter)
  entrezDEGs <- entrezDEGs$Entrez_ID
  if (length(entrezDEGs) < 10) {
    plotList[[i]] <- NA
    next
  }
  DEG_pathways <- clusterProfiler::enricher(entrezDEGs,
                                            universe = entrez_IDs,
                                            pAdjustMethod = "fdr",
                                            pvalueCutoff = 0.05, #p.adjust cutoff
                                            TERM2GENE = wpid2gene,
                                            TERM2NAME = wpid2name)
  DEG_pathways <- DOSE::setReadable(DEG_pathways, species_data$orgdb, keyType = "ENTREZID")
  ###################### TODO FIX THIS CHANTGED FOR TESTING ######################
  DEG_pathways@result$p.adjust[1:10] <- rep(0.001,10)
  DEG_pathways@result$qvalue[1:10] <- rep(0.001,10)
  ###################### TODO FIX THIS CHANTGED FOR TESTING ######################
  if (nrow(DEG_pathways@result %>% filter(p.adjust < 0.01 & qvalue < 0.05)) > 1) {
    plotList[[i]] <- clusterProfiler::dotplot(DEG_pathways, showCategory = 20, font.size = 9)  +
      theme(axis.text.y = element_text(angle = 0)) +
      scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 30))
    write.table(as.data.frame(DEG_pathways),
                file = file.path(paths$pathway_analysis,paste0("Enriched.WikiPathways.",
                                                         str_replace_all(levels(significantResults$contrast)[i],
                                                                         " ",
                                                                         "_"),
                                                         ".txt")))
  } else { plotList[[i]] <- NA }
}
```

```{r 'pathway_analysis_by_contrast_print_plots', results='asis', eval=params$run_pathway_analysis, include=params$run_pathway_analysis, warning = FALSE, message = FALSE}
for (i in seq_along(contrasts)) {
  if (any(!is.na(plotList[[i]]))) {
    cat("###", contrasts[i], "  \n\n")
    print(plotList[[i]])
    cat('  \n\n')
  }
}

```


## GSEA: By contrast {.tabset .tabset-fade}

This section uses clusterProfiler `gseKEGG()` to detect enriched KEGG pathways and enrichplot `gseaplot2` to plot running scores.

```{r 'gsea_analysis_by_contrast_make_plots', collapse=TRUE, eval=params$run_pathway_analysis, include=params$run_pathway_analysis,  warning = FALSE, message = FALSE}
plotListGSEA <- list()
plotListGSEA_escore <- list()
contrasts <- unique(allResults$contrast)
for (i in seq_along(contrasts)) {
  DEGs_full <- dplyr::left_join(allResults %>% filter(contrast ==  contrasts[i]),
                                id_table_entrez,
                                by = params$biomart_filter)
  foldChanges <- DEGs_full %>% dplyr::pull(log2FoldChange)
  names(foldChanges) <- DEGs_full %>% dplyr::pull(Entrez_ID)
  foldChanges <- foldChanges %>% sort() %>% rev()
  foldChanges <- foldChanges[!is.na(names(foldChanges))]
  if (length(foldChanges) < 10) {
    plotListGSEA[[i]] <- NA
    plotListGSEA_escore[[i]] <- NA
    next
  }
  kk <- gseKEGG(foldChanges, keyType = "ncbi-geneid", organism = species_data$kegg_organism)
  gsea_sorted <- kk %>% arrange(desc(enrichmentScore))
  gsea_sig <- gsea_sorted %>% filter(p.adjust < 0.05)
  write.table(as.data.frame(gsea_sorted),
              file = file.path(paths$pathway_analysis,paste0("Enriched.KEGG_Pathways.",
                                                       str_replace_all(contrasts[i]," ","_"),
                                                       ".txt")))
  if (dim(kk@result)[1] > 1) {
    plotListGSEA[[i]] <- ridgeplot(kk)
    plotListGSEA_escore[[i]] <- enrichplot::gseaplot2(gsea_sig, geneSetID = c(seq_along(gsea_sig$ID)))
  } else {
    plotListGSEA[[i]] <- NA
    plotListGSEA_escore[[i]] <- NA
  }
}
```

```{r 'gsea_analysis_by_contrast_print_plots', results='asis', eval=params$run_pathway_analysis, include=params$run_pathway_analysis,  warning = FALSE, message = FALSE}
for (i in seq_along(contrasts)) {
  if (any(!is.na(plotListGSEA[[i]]))) {
    cat("###", contrasts[i], "  \n\n")
    print(plotListGSEA[[i]])
    cat('  \n\n')
  }
}
```

## GSEA Enrichment Score Plots: By contrast {.tabset .tabset-fade}

In this section, only the pathway with the top enrichment score is shown (for simplicity). Note that it is possible to plot any pathway of interest by altering the call to `gseaplot`.

```{r 'gsea_scores_plot', results='asis', eval=params$run_pathway_analysis, include=params$run_pathway_analysis,  warning = FALSE, message = FALSE}
for (i in seq_along(contrasts)) {
  if (any(!is.na(plotListGSEA_escore[[i]]))) {
    cat("###", contrasts[i], "  \n\n")
    print(plotListGSEA_escore[[i]])
    cat('  \n\n')
  }
}
```
# Session Info

## Date the report was generated.

```{r reproducibility1, echo = FALSE}
## Date the report was generated
Sys.time()
```

## Version of R-ODAF Repository

The git hash for the commit of the [R-ODAF repository](`r system("git remote get-url origin", intern = T)`) used in this analysis is `r system("git rev-parse HEAD", intern = T)`.

## Parameters Used

From the list elements in each params${variable} used to generate this report.

```{r paramsList, echo = FALSE}
df <- as.data.frame(unlist(params))
names(df) <- "Parameter Value"
knitr::kable(as.data.frame(df), format = "markdown")
```

## Wallclock time spent generating the report.

```{r reproducibility2, echo = FALSE}
## Processing time in seconds
totalTime <- diff(c(startTime, Sys.time()))
round(totalTime, digits = 3)
```

## `R` session information.

```{r reproducibility3, echo = FALSE}
## Session info
options(width = 120)
session_info()
```

## Pandoc version used: `r rmarkdown::pandoc_version()`.