
# Session Info {.tabset}

## Date the report was generated.

```{r reproducibility1, echo = FALSE}
## Date the report was generated
Sys.time()
```

## Version of R-ODAF Repository

The git hash for the commit of the [R-ODAF repository](`r system("git remote get-url origin", intern = T)`) used in this analysis is `r system("git rev-parse HEAD", intern = T)`.

## Parameters Used

From the list elements in each params${variable} used to generate this report.

```{r paramsList, echo = FALSE}
pars <- params # otherwise it's read-only
pars$biospyder <- NULL # remove the manifest from the params
df <- as.data.frame(unlist(pars))
names(df) <- "Parameter Value"
knitr::kable(as.data.frame(df), format = "markdown")
```

## Wallclock time spent generating the report

```{r reproducibility2, echo = FALSE}
## Processing time in seconds
totalTime <- diff(c(startTime, Sys.time()))
round(totalTime, digits = 3)
```

## `R` session information

```{r reproducibility3, echo = FALSE}
options(width = 120)
sessioninfo::session_info()

```

## Pandoc version

Pandoc version used: `r rmarkdown::pandoc_version()`.
